Does freshwater connectivity influence phosphorus retention in lakes?

### 2019-01-22 manuscript/manuscript.Rmd
towards including tables
add citations

@@

We modelled lake total phosphorus retention (hereafter, P retention) using the Vollenweider equation that models P retention as a function of water residence time and a parameter (k) conceptually representing in-lake P decay (Chapra 2008, Vollenweider 1975). Although there are several variants of this basic equation, we selected a 2-parameter form (equation 1) that has been
shown to have good performance in multiple cross-sectional studies (Brett and Benjamin 2007, Cheng et al. 2010):

@@

where $R_i$ is P retention as a fraction of P inputs, i $\tau$ is water residence time, k is a unitless parameter representing in-lake P decay, and w
 x is a unitless parameter representing P export via hydrologic flushing. Here, higher values of k are associated with greater integrated P losses from sedimentation and biological uptake resulting in greater P retention. Note that Eq 1 does not include a recycling term. Therefore, our results represent net l P retention (as opposed to gross P retention) under a steady state assumption where lakes are at equilibrium with respect to recycling (Vollenweider 1975). Note that although some forms of the Vollenweider equation use P loading as a predictor variable, it does not appear in Eq 1. The reason for this is two-fold. First, estimates of P loading are more difficult to obtain than estimates of water residence time and our aim was to develop a model than can be widely applied to lakes for which we lack detailed data on P loading. Second, loading based model forms have been shown to be mathematically equivalent to water residence time based model forms (Brett and Benjamin 2008).
We used the model described by Eq 1 to compare lake P retention in lakes with different connectivity by fitting the model in two ways. First, we modelled the overall relationship between P retention and water residence time for all lakes in our dataset (global model). Second, we fit hierarchical versions of Eq. 1 where k was modelled separately (kj) as a function of a binary sub-population indicator gi denoting membership in one of two lake classes formed on the
basis of specific connectivity metrics (or lake depth) and specific spatial extents:

@@

where greater differences in k between R the two groups indicate greater support for a connectivity effect on P retention. Prior to model fitting, we examined the bivariate relationships between each connectivity metric, water residence i time, P loading, and other factors related to P retention
eusing Pearson’s correlation coefficients. The purpose of this exercise was twofold, to determine the potential for collinearity among any of the variables in Eq 2 and to identify any relationships between P retention, water residence time, and other watershed and lake characteristics that were not included in our model. As only one connectivity metric was used to define g for each model we did not use the results of this exercise to exclude variables from further investigation. We quantified the relative support for an effect of each connectivity metric on P retention in more detail by calculating the difference in the median value of the P decay parameter k between groups (i.e. $\Delta$k). We used these median k values along with median estimates of x to determine how differences in k translated to differences in P retention (Eq 2). We judged significance by whether or not differences in group-wise P retention were greater than the measurement precision of P retention (> 0.01).

We fit all models in a Bayesian framework using the non-linear extension to the brms package to interface with the Stan statistical program (Burkner 2017, Stan Development Team 2017). In both models, we set a semi-informative prior on k and x of N(1.3, 0.1) and N(0.45, 0.1) respectively. These priors were based on the confidence intervals presented in Brett and Benjamin (2007) and qualitatively matched those used by Cheng et al. (2010). We used the default settings of brms and rstan to generate posterior estimates using four chains of 4,000 iterations each with no thinning and initial parameter values drawn from a uniform distribution bounded between -2 and 2. We also used the brms package for model evaluation by computing a Bayesian $R^{2}$ following the method of Gelman et al. (2017).

@@

Our lake connectivity classes were formed by dividing the lake dataset into two classes for each connectivity metric based on the bivariate relationship between each metric and P retention. Prior studies have used a similar binary splitting approach to examine the effect of various exogenous factors on lake P retention (Cheng et al. 2010; Shimoda and Arhonditsis 2015). We determined splitting criteria for each metric from the results of a random forest procedure incorporating conditional inference trees (ctree, Hothorn and Zeileis 2015). This procedure creates binary splits of the independent variables (i.e. each of the connectivity metrics), which are recursively repeated to find the split that maximizes association with the dependent variable (P retention). The advantage of the ctree technique over the more typical classification-regression tree (CART) technique is that tree growth stopping rules are pre-specified (Hothorn et al., 2006). As a result, some of subjectivity associated with post-hoc tree pruning is avoided.

@@

A secondary purpose of examining the bivariate relationships between connectivity, water residence time, P loading and other factors related to P retention was to determine if our connectivity metrics were related to each other such that they provide similar information. With the exception of stream density and baseflow (r = -0.52, p < 0.05), correlations among connectivity metrics were low and of a similar magnitude as the correlations between connectivity metrics and water residence time (0.10 < r < 0.29). The strongest correlation between any connectivity metric or lake characteristic was upstream lake area and lake watershed area (r = 0.77, p < 0.05).
We observed notably strong correlations between P loading and several lake characteristics not accounted for in Eq 2. These include the correlation between P loading and lake watershed area (r = 0.62, p < 0.05) as well as the correlation between P loading and lake subwatershed area (r = o
 0.76, p < 0.05). In addition, the correlation between P loading and upstream lake area was quite strong (r = 0.52, p < 0.05). Despite strong correlations between P loading, watershed area, and upstream lake area, we did not observe a strong correlation between P loading and P retention (r = 0.15, p > 0.05). Notably, this correlation was much weaker than the correlation between water residence time and P retention (r = 0.44, p > 0.05). Our observation that the correlation between P retention and P loading was not appreciably stronger
than correlations between P retention and our connectivity metrics suggests that our estimates of connectivity metric effects on P retention are not confounded by interactions between water residence time and P loading.
Although we did not observe strong correlations among connectivity metrics, we found that lakes with similar connectivity metric values, were spatially clustered (Figure 5). In particular, we found that lakes were concentrated in either the southern or northern portions of our study area depending on their connectivity metric value (Figure 5). This observation is consistent with the findings of Fergus et al. (2017) that lakes in the northern portion of our study area have distinct freshwater connectivity as compared to lakes in the southern portion of our study area.

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.6\textwidth,keepaspectratio]{../figures/08_maps-1}
  \end{center}
  \caption{Locations of lakes of with different connectivity metric values at the (A) lake subwatershed and (B) lake watershed extent. Low valued lakes are represented by lighter (green) symbols, and high values lakes are represented by darker (purple) symbols. WS is the lake watershed extent. SWS is the lake subwatershed extent. Connectivity metrics are defined in Figure 2.}\label{fig5}
\end{figure*}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.5\textwidth,keepaspectratio]{../figures/04_global_vollenweider-1}
  \end{center}
  \caption{Residence time (yr) versus P retention for the NES dataset and the global model fit to the data ($R^{2}$ = 0.34, n = 129) where the solid line and shaded interval represents the median and central 95% interval estimates respectively (A). As above except that solid lines and shaded interval estimates represent hierarchical model fits to the data ($R^{2}$ = 0.41, n = 129) based on watershed average link length, which had the strongest association of any connectivity metric relative to P retention (B). Equations for these lines at median water residence time for low and high link length lakes are: Rp = 1 - (1 / (1 + 1.32 $\protect\tau^{0.4}$)) and Rp = 1 - (1 / (1 + 1.08 $\protect\tau^{0.4}$)) respectively.}\label{fig6}
\end{figure}

@@

Differences in P retention among lakes with different connectivity metric values was reflected in differences among connectivity class-specific values of the P decay parameter k (Figure 7). The connectivity metric that had the most effect on k was average link length (Table 2). For instance, hierarchical models fit with the average link length metric had a median effect size of 0.23 and 0.05 for k and P retention respectively (Table 2), which means that for this metric, lakes with shorter average link lengths retained 4.7 to 4.9 % more P than lakes with longer average link lengths. The influence of lake-stream-based connectivity metrics on lake P retention was similar
to stream-based connectivity metrics (Figure 7). This suggests that both lake-stream-based connectivity between lakes and streams as well as stream-based connectivity of lakes and their terrestrial watersheds influence P retention (Figure 2).

@@

Although prior studies have found that lake P retention is related to water residence time (Brett and Benjamin, Cheng, et al. 2010, Vollenweider 1975), there is substantial variation around this relationship, particularly at intermediate water residence times. We found that some of this remaining variation could be explained using a hierarchical modelling framework that accounts for differences in freshwater connectivity among lakes. Although we found that the magnitude of this effect depends on the specific connectivity metric, our results are consistent with the findings of previous studies showing that connectivity metrics are associated with differences in lake carbon input fluxes and differences in lake nitrogen output fluxes (Cardille et al. 2007; Schmadel et al. 2018). We also found important differences in the association between connectivity metrics and P retention at different measurement extents. Specifically, we found that P retention was more strongly associated with connectivity measured at the broader lake watershed extent rather than connectivity measured at the finer lake subwatershed extent.

@@

We found that differences in P retention among lake connectivity classes was influenced by specific connectivity metrics including average link length, closest lake distance, and stream density (Table 2, Figure 7). Indeed, these metrics were more strongly associated with P retention than covariates typically used in statistical P retention models (e.g. lake depth, Figure 1). Note that we were able to examine the influence of specific metrics as separate effects apart from water residence time because our model structure treats them as hierarchical coefficients on the P decay parameter k. As a result, although water residence time remains the dominant effect on lake P retention, we were able to estimate the specific influence of each metric in a way that is not possible using integrated water residence time and connectivity metrics such as watershed transport capacity (Fraterrigo and Downing 2008).
Several of the connectivity metrics that were most strongly associated with P retention were weakly correlated with watershed size (Figure 4). The weak nature of these correlations are consistent with the mixed results of prior studies linking watershed size to lake processes. For example, Zimmer and McGlynn (2018) found that carbon export was related to watershed size,
but Smith et al. (2003) found that the nitrogen flux was not strongly related to watershed size.
Taken together, our results and the results of prior studies suggest that watershed size and lake depth alone may not always reflect functional differences in potential material transport. One consequence of a correlation between connectivity metrics and watershed size is that metrics derived from watershed size, such as catchment to lake area ratio, are also likely to be associated with connectivity. Catchment to lake area ratio in particular has been previously used as an approximation or proxy of water residence w
 time (Rasmussen et al. 1989; Sobek et al. 2007).

Although our results differ from those of Soranno et al. (2015) who found that the presence of an upstream lake connection was not strongly associated with catchment to lake area ratio, our results suggest that catchment to lake area ratio likely incorporates some connectivity information and caution is needed before using it as a proxy for water residence time.
Another lake characteristic that was strongly related to watershed size was P loading. In particular, positive correlations between P loading and watershed size are consistent with the idea that lakes in larger watersheds receive greater P loading (Prairie and Kalff 1986). A related observation is that the correlation between P loading and upstream lake area was positive. This can be explained by the fact that larger watersheds often have greater numbers of lakes nested within them (Zhang et al. 2012). Another notable result is that we observed a weaker correlation between P loading and watershed area relative to the correlation between P loading and subwatershed area. One explanation for this result is that land-use cover at the finer lake subwatershed extent has more of an influence on P loading than land-use cover at the broader lake watershed extent (Soranno et al. 2015).

@@

One of the challenges in quantifying the effect of freshwater connectivity on lake P retention is that it is not always known what spatial extent of the watershed is functionally connected to a lake. We found that connectivity at the broader lake watershed extent rather than connectivity at the finer lake subwatershed extent was more strongly associated with differences in lake P retention. In addition, we found that individual characteristics such as lake depth (or more discrete measures of connectivity such as the presence of an upstream lake) were more strongly associated with P retention than any of the other connectivity metrics measured at the lake subwatershed extent (Table 2). These findings contrast with Soranno et al. (2015) who examined the association between lake nutrient concentrations (as opposed to retention) and land-use measured at varying spatial extents and found that measurements at the finer lake subwatershed extent rather than measurements at the broader lake watershed extent were more strongly associated with lake nutrient concentrations. One explanation for the difference between our results and those of Soranno et al. (2015) is that connectivity metrics may reflect long-range watershed processes to a greater degree than land-use cover. An alternative explanation is that controls of lake P retention may differ compared to controls on lake P concentration.

@@

Prior studies at regional extents have shown that P retention in streams and rivers is largely determined by the fate of the particulate load fraction (Cushing et al. 1994; Kronvang 1992; Russell et al. 1998; Vanni et al. 2001). For instance, the findings of Jarvie et al. (2011) show that riverine P loads can be controlled by nonpoint-source P delivery of particulate P. Therefore, it stands to reason that P retention in lakes may also be largely determined by the fate of the particulate load fraction. However, in the context of our broad-scale study it is difficult to examine this relationship because although we have estimates of nonpoint versus point source loading we do not have explicit estimates of particulate load fraction for large numbers of lakes. Unfortunately, we cannot use nonpoint source loading as a direct proxy for particulate load fraction because the two quantities do not have a consistent relationship. For example, Russell et al. (1998) report that the particulate phosphorus fraction of nonpoint source loads can be anywhere between 62 - 90% while Djodjic and Markensten (2018) report that this fraction can be anywhere between 33 and 80%. This may explain why prior broad scale studies that estimate lake P retention have not attempted to estimate separate effects of particulate versus dissolved loading (Alexander et al. 2008, Brett and Benjamin 2008).

We developed a conceptual model that places particulate load fraction in context with other processes affecting P retention (Figure 1). We expected that both land-stream and stream-lake connectivity influences how much particulate P is transported into lakes, which in turn affects their P retention. This expectation is supported by the findings of Cushing et al. (1994) as well as Guy et al. (1994) showing that particulate P can be transported beyond the direct drainage from stream-adjacent hillslopes.

Despite our inability to test such differential transport processes across many lake watersheds at broad scales, we note that such processes are indirectly supported by our finding that connectivity metrics are associated with lake P retention. For example, differential transport of particulate matter whereby barriers to flux and differences in drainage path configuration only become apparent beyond fine spatial extents may explain why we observed stronger association of P retention with metrics measured at the broader lake watershed extent rather than metrics measured at the finer scale lake subwatershed extent. Another finding consistent with differential transport of particulate P is our observation that average link length, which approximates stream network structure and along-stream transport potential (Barbera and Rosso 1989), was one of the more strongly associated metrics with lake P retention. Finally, it is notable that our stream density metric was influential at thee lake watershed extent but not at the lake subwatershed extent. Given that the stream density metric captures the average distance or drainage potential between any streams in the network and their adjacent hillslopes, floodplains, and wetlands (see Leibowitz et al. 2018), this suggests that differences in terrestrial runoff of particulate matter from hillslope and riparian areas are likely to be important for P retention.

Taken together, our findings are consistent with the idea that both connectivity between lakes and streams as well as connectivity of lakes and their terrestrial watersheds affect lake P retention. This conclusion matches that of prior studies showing that aquatic transport of phosphorus and nitrogen at the sub-continental scale is strongly controlled by processes affecting along-stream flux such as reservoir trapping (Alexander et al. 2008, Schmadel et al. 2018).

>>>>

  \begin{center}
    \includegraphics[width=0.5\textwidth,keepaspectratio]{../tables/crop/01_lake_characteristics_table} \label{table1}
  \end{center}

@@

We modelled lake total phosphorus retention (hereafter, P retention) using the Vollenweider equation that models P retention as a function of water residence time and a parameter (k) conceptually representing in-lake P decay \citep{chapra2008surface, vollenweiderInputoutputModels1975}. Although there are several variants of this basic equation, we selected a 2-parameter form (equation 1) that has been
shown to have good performance in multiple cross-sectional studies \citep{brett2008, cheng2010}:

@@

where $R_i$ is P retention as a fraction of P inputs, i $\tau$ is water residence time, $k$ is a unitless parameter representing in-lake P decay, and $x$ is a unitless parameter representing P export via hydrologic flushing. Here, higher values of k are associated with greater integrated P losses from sedimentation and biological uptake resulting in greater P retention. Note that Eq 1 does not include a recycling term. Therefore, our results represent net P retention (as opposed to gross P retention) under a steady state assumption where lakes are at equilibrium with respect to recycling \citep{vollenweiderInputoutputModels1975}. Note that although some forms of the Vollenweider equation use P loading as a predictor variable, it does not appear in Eq 1. The reason for this is two-fold. First, estimates of P loading are more difficult to obtain than estimates of water residence time and our aim was to develop a model than can be widely applied to lakes for which we lack detailed data on P loading. Second, loading based model forms have been shown to be mathematically equivalent to water residence time based model forms \citep{brett2008}.
We used the model described by Eq 1 to compare lake P retention in lakes with different connectivity by fitting the model in two ways. First, we modelled the overall relationship between P retention and water residence time for all lakes in our dataset (global model). Second, we fit hierarchical versions of Eq. 1 where $k$ was modelled separately ($k_{j}$) as a function of a binary sub-population indicator $g_{i}$ denoting membership in one of two lake classes formed on the basis of specific connectivity metrics (or lake depth) and specific spatial extents:

@@

where greater differences in $k$ between R the two groups indicate greater support for a connectivity effect on P retention. Prior to model fitting, we examined the bivariate relationships between each connectivity metric, water residence time, P loading, and other factors related to P retention using Pearson’s correlation coefficients. The purpose of this exercise was twofold, to determine the potential for collinearity among any of the variables in Eq 2 and to identify any relationships between P retention, water residence time, and other watershed and lake characteristics that were not included in our model. As only one connectivity metric was used to define g for each model we did not use the results of this exercise to exclude variables from further investigation. We quantified the relative support for an effect of each connectivity metric on P retention in more detail by calculating the difference in the median value of the P decay parameter k between groups (i.e. $\Delta$k). We used these median k values along with median estimates of x to determine how differences in k translated to differences in P retention (Eq 2). We judged significance by whether or not differences in group-wise P retention were greater than the measurement precision of P retention (> 0.01).

We fit all models in a Bayesian framework using the non-linear extension to the `brms` package to interface with the Stan statistical program \citep{burkner2017brms, stan2017}. In both models, we set a semi-informative prior on $k$ and $x$ of N(1.3, 0.1) and N(0.45, 0.1) respectively. These priors were based on the confidence intervals presented in \citet{brett2008} and qualitatively matched those used by \citet{cheng2010}. We used the default settings of brms and rstan to generate posterior estimates using four chains of 4,000 iterations each with no thinning and initial parameter values drawn from a uniform distribution bounded between -2 and 2. We also used the brms package for model evaluation by computing a Bayesian $R^{2}$ following the method of \citet{gelmanRsquaredBayesianRegression2017}.

@@

Our lake connectivity classes were formed by dividing the lake dataset into two classes for each connectivity metric based on the bivariate relationship between each metric and P retention. Prior studies have used a similar binary splitting approach to examine the effect of various exogenous factors on lake P retention \citep{cheng2010, shimodaIntegratingHierarchicalBayes2015}. We determined splitting criteria for each metric from the results of a random forest procedure incorporating conditional inference trees \citep[ctree,]{hothorn2015partykit}. This procedure creates binary splits of the independent variables (i.e. each of the connectivity metrics), which are recursively repeated to find the split that maximizes association with the dependent variable (P retention). The advantage of the ctree technique over the more typical classification-regression tree (CART) technique is that tree growth stopping rules are pre-specified \citep{hothorn2006unbiased}. As a result, some of subjectivity associated with post-hoc tree pruning is avoided.

@@

A secondary purpose of examining the bivariate relationships between connectivity, water residence time, P loading and other factors related to P retention was to determine if our connectivity metrics were related to each other such that they provide similar information. With the exception of stream density and baseflow (r = -0.52, p < 0.05), correlations among connectivity metrics were low and of a similar magnitude as the correlations between connectivity metrics and water residence time (0.10 < r < 0.29). The strongest correlation between any connectivity metric or lake characteristic was upstream lake area and lake watershed area (r = 0.77, p < 0.05).
We observed notably strong correlations between P loading and several lake characteristics not accounted for in Eq 2. These include the correlation between P loading and lake watershed area (r = 0.62, p < 0.05) as well as the correlation between P loading and lake subwatershed area (r = o
 0.76, p < 0.05). In addition, the correlation between P loading and upstream lake area was quite strong (r = 0.52, p < 0.05). Despite strong correlations between P loading, watershed area, and upstream lake area, we did not observe a strong correlation between P loading and P retention (r = 0.15, p > 0.05). Notably, this correlation was much weaker than the correlation between water residence time and P retention (r = 0.44, p > 0.05). Our observation that the correlation between P retention and P loading was not appreciably stronger than correlations between P retention and our connectivity metrics suggests that our estimates of connectivity metric effects on P retention are not confounded by interactions between water residence time and P loading.
Although we did not observe strong correlations among connectivity metrics, we found that lakes with similar connectivity metric values, were spatially clustered (Figure 5). In particular, we found that lakes were concentrated in either the southern or northern portions of our study area depending on their connectivity metric value (Figure 5). This observation is consistent with the findings of \citet{fergusFreshwaterLandscapeLake2017} that lakes in the northern portion of our study area have distinct freshwater connectivity as compared to lakes in the southern portion of our study area.

\begin{figure*}
  \begin{raggedleft}
    \includegraphics[width=0.6\textwidth,keepaspectratio]{../figures/08_maps-1}
  \end{raggedleft}
  \caption{Locations of lakes of with different connectivity metric values at the (A) lake subwatershed and (B) lake watershed extent. Low valued lakes are represented by lighter (green) symbols, and high values lakes are represented by darker (purple) symbols. WS is the lake watershed extent. SWS is the lake subwatershed extent. Connectivity metrics are defined in Figure 2.}\label{fig5}
\end{figure*}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.5\textwidth,keepaspectratio]{../figures/04_global_vollenweider-1}
  \end{center}
  \caption{Residence time (yr) versus P retention for the NES dataset and the global model fit to the data ($R^{2}$ = 0.34, n = 129) where the solid line and shaded interval represents the median and central 95\% interval estimates respectively (A). As above except that solid lines and shaded interval estimates represent hierarchical model fits to the data ($R^{2}$ = 0.41, n = 129) based on watershed average link length, which had the strongest association of any connectivity metric relative to P retention (B). Equations for these lines at median water residence time for low and high link length lakes are: Rp = 1 - (1 / (1 + 1.32 $\protect\tau^{0.4}$)) and Rp = 1 - (1 / (1 + 1.08 $\protect\tau^{0.4}$)) respectively.}\label{fig6}
\end{figure}

@@

Differences in P retention among lakes with different connectivity metric values was reflected in differences among connectivity class-specific values of the P decay parameter k (Figure 7). The connectivity metric that had the most effect on k was average link length (Table 2). For instance, hierarchical models fit with the average link length metric had a median effect size of 0.23 and 0.05 for k and P retention respectively (Table \ref{table2}{2}), which means that for this metric, lakes with shorter average link lengths retained 4.7 to 4.9 % more P than lakes with longer average link lengths. The influence of lake-stream-based connectivity metrics on lake P retention was similar
to stream-based connectivity metrics (Figure 7). This suggests that both lake-stream-based connectivity between lakes and streams as well as stream-based connectivity of lakes and their terrestrial watersheds influence P retention (Figure 2).

\begin{center}
  \includegraphics[width=0.5\textwidth,keepaspectratio]{../tables/crop/02_model_results_table} \label{table2}
\end{center}

@@

Although prior studies have found that lake P retention is related to water residence time \citep{brett2008, cheng2010, vollenweiderInputoutputModels1975}, there is substantial variation around this relationship, particularly at intermediate water residence times. We found that some of this remaining variation could be explained using a hierarchical modelling framework that accounts for differences in freshwater connectivity among lakes. Although we found that the magnitude of this effect depends on the specific connectivity metric, our results are consistent with the findings of previous studies showing that connectivity metrics are associated with differences in lake carbon input fluxes and differences in lake nitrogen output fluxes \citep{cardilleCarbonWaterCycling2007, schmadelThresholdsLakeReservoir2018}. We also found important differences in the association between connectivity metrics and P retention at different measurement extents. Specifically, we found that P retention was more strongly associated with connectivity measured at the broader lake watershed extent rather than connectivity measured at the finer lake subwatershed extent.

@@

We found that differences in P retention among lake connectivity classes was influenced by specific connectivity metrics including average link length, closest lake distance, and stream density (Table 2, Figure 7). Indeed, these metrics were more strongly associated with P retention than covariates typically used in statistical P retention models (e.g. lake depth, Figure 1). Note that we were able to examine the influence of specific metrics as separate effects apart from water residence time because our model structure treats them as hierarchical coefficients on the P decay parameter k. As a result, although water residence time remains the dominant effect on lake P retention, we were able to estimate the specific influence of each metric in a way that is not possible using integrated water residence time and connectivity metrics such as watershed transport capacity \citep{fraterrigoInfluenceLandUse2008}.
Several of the connectivity metrics that were most strongly associated with P retention were weakly correlated with watershed size (Figure 4). The weak nature of these correlations are consistent with the mixed results of prior studies linking watershed size to lake processes. For example, \citet{zimmerLateralVerticalLongitudinal2018} found that carbon export was related to watershed size, but \citet{smithEutrophicationFreshwaterCoastal2003} found that the nitrogen flux was not strongly related to watershed size.
Taken together, our results and the results of prior studies suggest that watershed size and lake depth alone may not always reflect functional differences in potential material transport. One consequence of a correlation between connectivity metrics and watershed size is that metrics derived from watershed size, such as catchment to lake area ratio, are also likely to be associated with connectivity. Catchment to lake area ratio in particular has been previously used as an approximation or proxy of water residence time \citep{rasmussenHumicContentLake1989, sobekPatternsRegulationDissolved2007}.

Although our results differ from those of \citet{soranno2015} who found that the presence of an upstream lake connection was not strongly associated with catchment to lake area ratio, our results suggest that catchment to lake area ratio likely incorporates some connectivity information and caution is needed before using it as a proxy for water residence time.
Another lake characteristic that was strongly related to watershed size was P loading. In particular, positive correlations between P loading and watershed size are consistent with the idea that lakes in larger watersheds receive greater P loading \citep{t-prairieEffectCatchmentSize1986}. A related observation is that the correlation between P loading and upstream lake area was positive. This can be explained by the fact that larger watersheds often have greater numbers of lakes nested within them (Zhang et al. 2012). Another notable result is that we observed a weaker correlation between P loading and watershed area relative to the correlation between P loading and subwatershed area. One explanation for this result is that land-use cover at the finer lake subwatershed extent has more of an influence on P loading than land-use cover at the broader lake watershed extent \citep{soranno2015}.

@@

One of the challenges in quantifying the effect of freshwater connectivity on lake P retention is that it is not always known what spatial extent of the watershed is functionally connected to a lake. We found that connectivity at the broader lake watershed extent rather than connectivity at the finer lake subwatershed extent was more strongly associated with differences in lake P retention. In addition, we found that individual characteristics such as lake depth (or more discrete measures of connectivity such as the presence of an upstream lake) were more strongly associated with P retention than any of the other connectivity metrics measured at the lake subwatershed extent (Table 2). These findings contrast with \citet{soranno2015} who examined the association between lake nutrient concentrations (as opposed to retention) and land-use measured at varying spatial extents and found that measurements at the finer lake subwatershed extent rather than measurements at the broader lake watershed extent were more strongly associated with lake nutrient concentrations. One explanation for the difference between our results and those of \citet{soranno2015} is that connectivity metrics may reflect long-range watershed processes to a greater degree than land-use cover. An alternative explanation is that controls of lake P retention may differ compared to controls on lake P concentration.

@@

Prior studies at regional extents have shown that P retention in streams and rivers is largely determined by the fate of the particulate load fraction \citep{cushingTransportDynamicsFine1993, kronvang1992, russell1998, vanni2001}. For instance, the findings of \citet{jarvie2011} show that riverine P loads can be controlled by nonpoint-source P delivery of particulate P. Therefore, it stands to reason that P retention in lakes may also be largely determined by the fate of the particulate load fraction. However, in the context of our broad-scale study it is difficult to examine this relationship because although we have estimates of nonpoint versus point source loading we do not have explicit estimates of particulate load fraction for large numbers of lakes. Unfortunately, we cannot use nonpoint source loading as a direct proxy for particulate load fraction because the two quantities do not have a consistent relationship. For example, \citet{russell1998} report that the particulate phosphorus fraction of nonpoint source loads can be anywhere between 62 - 90% while \citet{djodjicSingleFieldsRiver2018} report that this fraction can be anywhere between 33 and 80%. This may explain why prior broad scale studies that estimate lake P retention have not attempted to estimate separate effects of particulate versus dissolved loading \citep{alexander2008, brett2008}.

We developed a conceptual model that places particulate load fraction in context with other processes affecting P retention (Figure 1). We expected that both land-stream and stream-lake connectivity influences how much particulate P is transported into lakes, which in turn affects their P retention. This expectation is supported by the findings of \citet{cushingTransportDynamicsFine1993} as well as \citet{guy1994} showing that particulate P can be transported beyond the direct drainage from stream-adjacent hillslopes.

Despite our inability to test such differential transport processes across many lake watersheds at broad scales, we note that such processes are indirectly supported by our finding that connectivity metrics are associated with lake P retention. For example, differential transport of particulate matter whereby barriers to flux and differences in drainage path configuration only become apparent beyond fine spatial extents may explain why we observed stronger association of P retention with metrics measured at the broader lake watershed extent rather than metrics measured at the finer scale lake subwatershed extent. Another finding consistent with differential transport of particulate P is our observation that average link length, which approximates stream network structure and along-stream transport potential \citep{labarberaFractalDimensionStream1989}, was one of the more strongly associated metrics with lake P retention. Finally, it is notable that our stream density metric was influential at thee lake watershed extent but not at the lake subwatershed extent. Given that the stream density metric captures the average distance or drainage potential between any streams in the network and their adjacent hillslopes, floodplains, and wetlands \citep[see]{leibowitzConnectivityStreamsWetlands2018}, this suggests that differences in terrestrial runoff of particulate matter from hillslope and riparian areas are likely to be important for P retention.

Taken together, our findings are consistent with the idea that both connectivity between lakes and streams as well as connectivity of lakes and their terrestrial watersheds affect lake P retention. This conclusion matches that of prior studies showing that aquatic transport of phosphorus and nitrogen at the sub-continental scale is strongly controlled by processes affecting along-stream flux such as reservoir trapping \citep{alexander2008, schmadelThresholdsLakeReservoir2018}.
